---
title: "Ripped Foil - Experiment 2"
author: "Zeynep Civelek"
date: "June 30, 2020"
output: word_document
---

```{r, include=FALSE}
#PREPARE
R.Version()#for referencing, shows you which R version you are using
rm(list=ls())#removes any other items in your workspace
ls()#check whether workspace is empty
```

```{r, include=FALSE}
#Directory is set, data is added and all the libraries necessary for running the analyses are loaded.

ripmon1<-read.csv("ripped_monkey_test16.csv", header=T) #Analyses are done on the last 16 trials of Test and 16 trials of Transfer phase. This data file includes those only.
#all the libraries necessary to run the analyses
library(lme4)
library(readr)
library(tidyverse)
library(sjPlot)
#install.packages('TMB', type='source')
library(ggthemes)
library(gridExtra)
library(reshape2)
library(car)
#library(psych)
library("ggpubr")
library(dplyr)

```

##Exploring data 

The structure of the data is shown below with all the variable names and levels of measurement.

```{r, echo=FALSE}
ripmon1$id<-as.factor(ripmon1$id)
ripmon1$sex<-as.factor(ripmon1$sex)
ripmon1$trialtype<-as.factor(ripmon1$trialtype)
ripmon1$phase2<-as.factor(ripmon1$phase2)
ripmon1$phase<-as.factor(ripmon1$phase)

str(ripmon1)
summary(ripmon1)
sd(ripmon1$trialno)
sd(ripmon1$age)

```

Exploring the Warm-up (named warm1) (locating the reward when there is no foil covering the cups) and Pre-test (named warm2 in data) (locating the reward when the cups are covered with foil) phases.

```{r, echo=FALSE}
ripmon_complete<-read.csv("ripped_monkey_complete.csv", header=T)#This data contains the complete data.
ripmon_warm1 <- ripmon_complete %>%
  filter(sessiontype=="warmup") %>%
  group_by(id, sex, age, sessionno, trialno) %>% 
  summarize(correct)

ripmon_warm2 <- ripmon_complete %>%
  filter(sessiontype=="foil") %>%
  group_by(id, sex, age, sessionno, trialno) %>% 
  summarize(correct)

range(ripmon_warm1$sessionno)
range(ripmon_warm2$sessionno)
```
**In warm-up, monkeys got up to 18 sessions to reach criterion and in pre-test, up to 6 sessions (or they'd receive up to 10 sessions). The minimum number of sessions to reach criterion was 2 (14/16 correct)**

I aggregated the trial-by-trial data to create one score per monkey for Test and Transfer phases so I can see the distribution of scores.

```{r, echo=FALSE}
ripmon_individual <- ripmon1 %>%
  filter(!is.na(correct)) %>%
  group_by(id) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

ripmon_test <- ripmon1 %>%
   filter(!is.na(correct)) %>%
  filter(phase=="test") %>%
  group_by(id) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

ripmon_transfer <- ripmon1 %>%
   filter(!is.na(correct)) %>%
  filter(phase=="transfer") %>%
  group_by(id) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

ripmon_separate <- ripmon1 %>%
   filter(!is.na(correct)) %>%
  group_by(id, phase,age) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

```

##Histograms for Test and Transfer Phases and the overall score

```{r echo=FALSE}
hist(ripmon_test$correct, xlab="Mean score in Test", xlim=c(0,1))
hist(ripmon_transfer$correct, xlab="Mean score in Transfer", xlim=c(0,1))
hist(ripmon_individual$correct, xlab="Overall score (Test and Transfer)", xlim=c(0,1))
```

##Preparation of the data for running the GLMM

* Scaling age and trial number
* Coding categorical variables as dummy variables 
* Centering the slopes.

```{r, echo=FALSE}
#We scale variables so that the differences in the range of the variables do not influence the parameter estimations.
ripmon1$z.trialno<-as.vector(scale(ripmon1$trialno))
ripmon1$z.age=as.vector(scale(ripmon1$age))
ripmon1$trialtype<-relevel(ripmon1$trialtype, ref = "stick")

#coding dummy variables before centering the slopes
ripmon1$trialtype.food<-as.numeric(ripmon1$trialtype==levels(ripmon1$trialtype)[2])
ripmon1$sex.m<-as.numeric(ripmon1$sex==levels(ripmon1$sex)[2])
ripmon1$phase.transfer=as.numeric(ripmon1$phase==levels(ripmon1$phase)[2])

#centering the slopes: p-values of the factors can be influenced by the choice of reference category.
#by centering the factor for the random slope components the p-values should be the same irrespective of the choice of the reference level
ripmon1$trialtype.food.c<-ripmon1$trialtype.food-mean(ripmon1$trialtype.food)
ripmon1$phase.transfer.c=ripmon1$phase.transfer-mean(ripmon1$phase.transfer)
ripmon1$sex.m.c<-ripmon1$sex.m -mean(ripmon1$sex.m)
summary(ripmon1)

```

```{r, echo=FALSE}
source("./Roger_functions/diagnostic_fcns.r")
source("./Roger_functions/glmm_stability.r")
source("./Roger_functions/boot_glmm.r")
```


##Full model - GLMM 01 Experiment 2

```{r, echo=TRUE}
contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))
full=glmer(correct ~ trialtype*phase+z.trialno+z.age+sex+(1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
```

The full model does not include box type (the location of the blue/pink box) as a random effect anymore due to convergence issues. The correlations between random slope and random intercept are not included either.

##Model assumptions

####Distribution of random effects
```{r, echo=FALSE}
#An assumption of GLMMs is that the BLUPs are normally distributed. We use the following function to evaluate this. BLUPs are Best Linear Unbiased Predictors: the estimated grouping factor's level specific deviation from the common intercept or slope.
ranef.diagn.plot(full)
```

####Model stability

```{r, echo=FALSE}
#summary reveals min and max of the estimates obtained after casewise deletion of levels of random effects together with the original estimates.
# m.stab=glmm.model.stab(model.res=full, contr=contr)
# m.stab.plot(m.stab$summary[,-1])
# m.stab$summary
# m.stab$detailed$warnings
```

####Multicollinearity

```{r, echo=FALSE}
coll=lm(correct ~ trialtype+phase+z.trialno+z.age+sex, data=ripmon1)
round(vif(coll),3)
#no vif issues
```

#These indicate no issues.

##Null model - GLMM 01 Experiment 2

```{r, echo=TRUE}
null=glmer(correct~z.trialno+(1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
```
**Added trial no to the null model to control for its effects.**

##Full and Null comparison
```{r, echo=FALSE}
round(anova(null, full, test="Chisq"),3)
```
**The full model is significantly different than the null model.**

##Model output - GLMM 01 Experiment 2 (Table S8)

####Coefficients

```{r, echo=FALSE}
round(summary(full)$coefficients, 3)
```

####Individual predictor: Likelihood tests

```{r, echo=FALSE}
xdrop1=drop1(full, test="Chisq",control=contr)
round(xdrop1,3)
```
**There is a significant effect of sex. The interaction term is not significant.**


```{r, include=FALSE}
####Confidence intervals for the full model with the interaction

#ripmon_model1=boot.glmm.pred(model.res=full, excl.warnings=T, nboots=1000, para=T)
#save.image("ripmon_model1_CIs.RData")
load("ripmon_model1_CIs.RData")
round(ripmon_model1$ci.estimates, 3)
```

**I will remove the non-significant interaction term to reduce the model.**

##Reduced model without the interaction - GLMM 02 Experiment 2
##Full model 2

```{r, echo=TRUE}
full2=glmer(correct ~ trialtype+phase+z.trialno+z.age+sex+(1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
summary(full2)
```

**The model summary show that there is a significant effect of sex and trial type.**

##Null model 2

```{r, echo=TRUE}
null2<-glmer(correct~ z.trialno + (1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
```

##Full and null model 2 comparison

```{r, echo=FALSE}
round(anova(null2, full2, test="Chisq"),3)
```
**There is a significant difference between the full and the null model.**

##Model output - GLMM 02 Experiment 2 (Table S9)

####Coefficients

```{r, echo=FALSE}
round(summary(full2)$coefficients, 3)

```

####Individual predictors : Likelihood ratio tests
  
```{r echo=FALSE}
xdrop1=drop1(full2, test="Chisq",control=contr)
round(xdrop1,3)
```
**The trial type and sex are significant predictors of performance.**

```{r, include=FALSE}
####Confidence intervals for the reduced model without the interaction

#ripmon_model2=boot.glmm.pred(model.res=full2, excl.warnings=T, nboots=1000, para=T)
#save.image("ripmon_model2_CIs.RData")
load("ripmon_model2_CIs.RData")
round(ripmon_model2$ci.estimates, 3)
```

Overall, monkeys performed better in food-stick trials than in stick-food trials, not differently in Test and Transfer- (and males better than females).

##Test against chance
####Test and transfer phase

```{r, echo=TRUE}
testphase <-ripmon1%>%
  filter(phase == "test")
testchance <- glmer(correct ~ 1 + (z.trialno | id),data= testphase, family=binomial)
summary(testchance)
confint(object=testchance)
```

**The intercept is significant: Performance in Test phase is above chance level.**

```{r, echo=TRUE}
transferphase <-ripmon1%>%
  filter(phase == "transfer")
transferchance <- glmer(correct ~ 1 + (z.trialno | id),data= transferphase, family=binomial)
summary(transferchance)
```

**Performanc in Transfer phase is above chance level too.**

####Trial type

```{r, echo=FALSE}
sticktrials <-ripmon1%>%
  filter(trialtype == "stick")
stickchance <- glmer(correct ~ 1 + (z.trialno | id),data= sticktrials, family=binomial)
summary(stickchance)
```


```{r echo=FALSE}
foodtrials <-ripmon1%>%
  filter(trialtype == "food")
foodchance <- glmer(correct ~ 1 + (z.trialno | id),data= foodtrials, family=binomial)
summary(foodchance)
```
**Monkeys performed above chance levels in both stick-food and food-stick trials.**

##Correcting the p-values for the post-hoc tests using Hochberg correction (in the order: test, transfer, stick-food, food-stick)

```{r, echo=TRUE}
pvalues <- c(3.69e-06, 2.06e-08, 1.41e-05, 5.25e-09)
p.adjust(pvalues, method="hochberg")
```

##For comparison purposes with children, I run two additional tests
#### 1)First trial performance in Transfer phase: Is performance above chance?

```{r, echo=TRUE}
ripmon_1st_trial <- ripmon1 %>%
  filter(phase=="transfer" & sessionno=="1", trialno=="1")

sum(ripmon_1st_trial$correct)
length(ripmon_1st_trial$correct)

binom.test(x=sum(ripmon_1st_trial$correct), n=length(ripmon_1st_trial$correct), p=0.5, alternative = "two.sided")
```

**No, performance in the first trial of transfer phase does not differ from chance**

####2)Performance of monkeys in the first 8 trials of Test phase: Is it above chance level?

```{r, echo=FALSE}
ripmon_complete$z.trialno<-as.vector(scale(ripmon_complete$trialno))
ripmon_test8 <- ripmon_complete %>%
  filter(sessiontype=="test" & sessionno=="1")
test8chance<-glmer(correct ~ 1 + (z.trialno | id), data= ripmon_test8, family=binomial)
summary(test8chance)
```
**No, they are at chance level in the first 8 trials of Test phase.**


##Scatter plot of performance in Test and Transfer phases across all trials - Figure 5

####To do this scatter plot I will run two new models to calculate CIs for Test and Transfer so we can plot the performance in test and transfer across trial numbers separately.

```{r}
###First the Test phase
#getting data and doing transformation for the CI model.

ripmon_test<-read.csv("ripped_monkey_complete_test.csv", header=T)
str(ripmon_test)

ripmon_test$id<-as.factor(ripmon_test$id)
ripmon_test$sex<-as.factor(ripmon_test$sex)
ripmon_test$phase<-as.factor(ripmon_test$phase)
ripmon_test$sessiontype<-as.factor(ripmon_test$sessiontype)

ripmon_test$trialtype<-as.factor(ripmon_test$trialtype)


#We scale variables so that the differences in the range of the variables do not influence the parameter estimations.
ripmon_test$z.trialno<-as.vector(scale(ripmon_test$trialno))
ripmon_test$z.age=as.vector(scale(ripmon_test$age))
ripmon_test$trialtype<-relevel(ripmon_test$trialtype, ref = "stick")

#coding dummy variables before centering the slopes
ripmon_test$trialtype.food<-as.numeric(ripmon_test$trialtype==levels(ripmon_test$trialtype)[2])
ripmon_test$sex.m<-as.numeric(ripmon_test$sex==levels(ripmon_test$sex)[2])

#centering the slopes: p-values of the factors can be influenced by the choice of reference category.
#by centering the factor for the random slope components the p-values should be the same irrespective of the choice of the reference level
ripmon_test$trialtype.food.c<-ripmon_test$trialtype.food-mean(ripmon_test$trialtype.food)
ripmon_test$sex.m.c<-ripmon_test$sex.m -mean(ripmon_test$sex.m)

summary(ripmon_test)

```



```{r}
##Then I will run a the model for calculating CIs for the Test phase.

#contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))

#full.CI.int <- glmer(correct ~ trialtype.food.c + z.age + sex.m.c + z.trialno+ (1|id)+(0+z.trialno+trialtype.food.c|id), data=ripmon_test, family=binomial, control=contr)

#boot.res.int=boot.glmm.pred(model.res=full.CI.int, excl.warnings=T, nboots=1000, resol=80, level=0.95, use="z.trialno", n.cores="all-1", para=T)

#boot.res.int$ci.estimates
#boot.res.int$ci.predicted

#save.image("ripped_mon_complete_testCIs.RData")
load("ripped_mon_complete_testCIs.RData")

```

```{r}
#Then I plot the graph for Test phase across trials.
ripmon_separate2 <- ripmon_test %>%
  group_by(trialno) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(trialno)%>%
    add_count(correct)

plot.int=seq(from=min(ripmon_separate2$trialno),
             to=max(ripmon_separate2$trialno),length.out=80)

ci.data<-boot.res.int$ci.predicted

scatter.int2 <- ggplot(ci.data, aes(y = fitted)) +
  geom_point(data=ripmon_separate2, aes(x = trialno, y = correct), size = ripmon_separate2$n, alpha = 0.8) +
  geom_ribbon(data = ci.data, aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data, aes(x = plot.int, y = fitted), lty = 2)+ scale_x_continuous(breaks = c(1,8,16,24,32,40,48,56,64,72,80)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Number of trials in Test") +
  ylab("Correct response") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.int2
```

```{r}
###Then I will do the same for Transfer phase.

ripmon_tra<-read.csv("ripped_monkey_complete_transfer.csv", header=T)# this is the data for the transfer phase only
str(ripmon_tra)

ripmon_tra$id<-as.factor(ripmon_tra$id)
ripmon_tra$sex<-as.factor(ripmon_tra$sex)
ripmon_tra$sessiontype<-as.factor(ripmon_tra$sessiontype)

ripmon_tra$trialtype<-as.factor(ripmon_tra$trialtype)


#We scale variables so that the differences in the range of the variables do not influence the parameter estimations.
ripmon_tra$z.trialno<-as.vector(scale(ripmon_tra$trialno))
ripmon_tra$trialtype<-relevel(ripmon_tra$trialtype, ref = "stick")
ripmon_tra$z.age=as.vector(scale(ripmon_tra$age))

#coding dummy variables before centering the slopes
ripmon_tra$trialtype.food<-as.numeric(ripmon_tra$trialtype==levels(ripmon_tra$trialtype)[2])
ripmon_tra$sex.m<-as.numeric(ripmon_tra$sex==levels(ripmon_tra$sex)[2])

#centering the slopes: p-values of the factors can be influenced by the choice of reference category.
#by centering the factor for the random slope components the p-values should be the same irrespective of the choice of the reference level
ripmon_tra$trialtype.food.c<-ripmon_tra$trialtype.food-mean(ripmon_tra$trialtype.food)
ripmon_tra$sex.m.c<-ripmon_tra$sex.m -mean(ripmon_tra$sex.m)

summary(ripmon_tra)

```

```{r}
##The model to calculate CIs for Transfer phase.

#contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))

#full.CI.int2 <- glmer(correct ~ trialtype.food.c + z.age + sex.m.c + z.trialno+ (1|id)+(0+z.trialno+trialtype.food.c|id), data=ripmon_tra, family=binomial, control=contr)

#boot.res.int2=boot.glmm.pred(model.res=full.CI.int2, excl.warnings=T, nboots=1000, resol=16, level=0.95, use="z.trialno", n.cores="all-1", para=T)

#boot.res.int2$ci.estimates
#boot.res.int2$ci.predicted

#save.image("ripped_mon_complete_transferCIs.RData")
load("ripped_mon_complete_transferCIs.RData")
```


```{r}
ripmon_separate3 <- ripmon_tra %>%
  group_by(trialno) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(trialno)%>%
    add_count(correct)

plot.int2=seq(from=min(ripmon_separate3$trialno),
             to=max(ripmon_separate3$trialno),length.out=16)

ci.data2<-boot.res.int2$ci.predicted

scatter.int3 <- ggplot(ci.data2, aes(y = fitted)) +
  geom_point(data=ripmon_separate3, aes(x = trialno, y = correct), size = ripmon_separate3$n, alpha = 0.8) +
  geom_ribbon(data = ci.data2, aes(x = plot.int2, y = fitted, ymin = lower.cl, ymax = upper.cl), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data2, aes(x = plot.int2, y = fitted), lty = 2)+ scale_x_continuous(breaks =c(1,8,16)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Number of trials in Transfer") +
  ylab("Correct response") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.int3

```


####Figre 5 with combining the two graphs
```{r}
ggarrange(scatter.int2, scatter.int3, widths= c(7,7), heights = c(10, 10),
          ncol = 2, nrow = 1)
```

