---
title: "Ripped Foil Experiment 1 - Children"
author: "Zeynep Civelek"
date: "June 10, 2020"
output: word_document
---

```{r, include=FALSE}
#PREPARE
R.Version()#for referencing, shows you which R version you are using
rm(list=ls())#removes any other items in your workspace
ls()#check whether workspace is empty
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Directory is set, data is added and all the libraries necessary for running the analyses are loaded.

ripchild<-read.csv("ripped_child.csv", header=T)
#all the libraries necessary to run the analyses
library(lme4)
library(readr)
library(tidyverse)
library(sjPlot)
#install.packages('TMB', type='source')
library(ggthemes)
library(gridExtra)
library(reshape2)
library(car)
#library(psych)
library("ggpubr")
library(dplyr)

```

##Exploring the data

The structure of the data is shown below with all the variable names and levels of measurement.

```{r echo=FALSE}
ripchild$age <- as.factor(ripchild$age)# makes continuous variables categorical
ripchild$id<-as.factor(ripchild$id)
ripchild$sex<-as.factor(ripchild$sex)
ripchild$trialtype<-as.factor(ripchild$trialtype)
ripchild$order<-as.factor(ripchild$order)
ripchild$order2<-as.factor(ripchild$order2)
ripchild$order3<-as.factor(ripchild$order3)
ripchild$agecat<-as.factor(ripchild$agecat)
ripchild$openended<-as.factor(ripchild$openended)

str(ripchild)
summary(ripchild)
mean(ripchild$agemonths)
sd(ripchild$agemonths)
sd(ripchild$trialno)
```

```{r}
###Getting the age, sex distribution table
ripchild_penfirst <- ripchild %>%
  filter(order=="penfirst") %>%
  group_by(id, sex, agemonths) %>% 
  summarize(correct = mean(correct))
boys_pen <- subset(ripchild_penfirst, sex == "2")
mean(boys_pen$agemonths)
sd(boys_pen$agemonths)
min(boys_pen$agemonths)
max(boys_pen$agemonths)
girls_pen <- subset(ripchild_penfirst, sex == "1")
mean(girls_pen$agemonths)
sd(girls_pen$agemonths)
min(girls_pen$agemonths)
max(girls_pen$agemonths)


ripchild_stickerfirst <- ripchild %>%
  filter(order=="stickerfirst") %>%
  group_by(id, sex, agemonths) %>% 
  summarize(correct = mean(correct))
boys_sticker <- subset(ripchild_stickerfirst, sex == "2")
mean(boys_sticker$agemonths)
sd(boys_sticker$agemonths)
min(boys_sticker$agemonths)
max(boys_sticker$agemonths)
girls_sticker <- subset(ripchild_stickerfirst, sex == "1")
mean(girls_sticker$agemonths)
sd(girls_sticker$agemonths)
min(girls_sticker$agemonths)
max(girls_sticker$agemonths)
```

The distribution of age:

```{r pressure, echo=FALSE}
hist(ripchild$agemonths, xlab="Age in months (36-72)", ylab="Frequency")
```

I aggregated the trial-by-trial data to create one score per child for Test and Transfer phases as well as an overall score so I can check the distributions of scores.

```{r echo=FALSE}
ripchild_individual <- ripchild %>%
  group_by(id, age, agemonths) %>% 
  summarize(correct = mean(correct))%>%
  ungroup()%>%
  group_by(agemonths)%>%
    add_count(correct)

mean(ripchild_individual$agemonths)
sd(ripchild_individual$agemonths)

ripchild_test <- ripchild %>%
  filter(order3=="test") %>%
  group_by(id, agemonths) %>% 
  summarize(correct = mean(correct))

ripchild_transfer <- ripchild %>%
  filter(order3=="transfer") %>%
  group_by(id, agemonths) %>% 
  summarize(correct = mean(correct)) 

```

##Histograms for Test and Transfer Phases and an overall score (ripchild_individual)

```{r echo=FALSE}
hist(ripchild_test$correct, xlab="Mean score in Test", xlim=c(0,1))
hist(ripchild_transfer$correct, xlab="Mean score in Transfer", xlim=c(0,1))
hist(ripchild_individual$correct, xlab="Overall score (Test and Transfer)", xlim=c(0,1))
```

##Preparation of the data for running the GLMM analysis

*Z-transformations for 'trial no' and 'age in months'
*Coding dummy variables for trialtype, sex and phase (order3), openended
*Centering the random slope components for these variables.

```{r include=FALSE}
#Preprocessing of data, Z-transforming trial no and age

ripchild$z.trialno<-as.vector(scale(ripchild$trialno))
ripchild$z.age=as.vector(scale(ripchild$agemonths))

ripchild$trialtype<-relevel(ripchild$trialtype, ref = "pen")
summary(ripchild)

```

```{r include=FALSE}
#Coding dummy variables before centering the slopes
ripchild$trialtype.sticker<-as.numeric(ripchild$trialtype==levels(ripchild$trialtype)[2])
ripchild$sex.m<-as.numeric(ripchild$sex==levels(ripchild$sex)[2])
ripchild$order3.transfer<-as.numeric(ripchild$order3==levels(ripchild$order3)[2])
ripchild$openended1<-as.numeric(ripchild$openended==levels(ripchild$openended)[2])


#Centering the slopes
ripchild$trialtype.sticker.c<-ripchild$trialtype.sticker-mean(ripchild$trialtype.sticker)
ripchild$sex.m.c<-ripchild$sex.m -mean(ripchild$sex.m)
ripchild$order3.transfer.c<-ripchild$order3.transfer-mean(ripchild$order3.transfer)
ripchild$openended1.c<-ripchild$openended1 -mean(ripchild$openended1)

```


```{r echo=FALSE}
source("./Roger_functions/diagnostic_fcns.r")
source("./Roger_functions/glmm_stability.r")
source("./Roger_functions/boot_glmm.r")
```

##Tests against chance
####Test and transfer phase

```{r, echo=TRUE}
##We first explored performance against chance level in Test and Transfer phases.
testphase <-ripchild%>%
  filter(order3 == "test")
testchance <- glmer(correct ~ 1 + (z.trialno | id),data= testphase, family=binomial)
summary(testchance)

transferphase <-ripchild%>%
  filter(order3 == "transfer")
transferchance <- glmer(correct ~ 1 + (z.trialno | id),data= transferphase, family=binomial)
summary(transferchance)
```

**The intercepts are significant: Performances in Test and Tansfer phases are above chance level.**

##Full model GLMM 01 Experiment 1 in ESM

```{r echo=TRUE}
contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))
full<-glmer(correct ~ trialtype*order3+z.trialno+z.age*order3+sex+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

The full model does not include box type (the location of the blue/pink box) as a random effect anymore due to convergence issues. But there is a singular fit warning here.

##Model assumptions

####Distribution of random effects
```{r echo=FALSE}
#An assumption of GLMMs is that the BLUPs are normally distributed. We use the following function to evaluate this. BLUPs are Best Linear Unbiased Predictors: the estimated grouping factor's level specific deviation from the common intercept or slope.
ranef.diagn.plot(full)
```

####Model stability

```{r echo=FALSE}
#summary reveals min and max of the estimates obtained after casewise deletion of levels of random effects together with the original estimates.
m.stab=glmm.model.stab(model.res=full, contr=contr)
m.stab.plot(m.stab$summary[,-1])
m.stab$summary
m.stab$detailed$warnings
```

####Multicollinearity

```{r echo=FALSE}
coll=lm(correct ~ trialtype+order3+z.trialno+z.age+sex, data=ripchild)
round(vif(coll),3)
#no vif issues
```

None of these indicate a problem.

##Null model GLMM 01 Experiment 1 in ESM

```{r echo=TRUE}
null<-glmer(correct~ z.trialno+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

##Full and Null comparison GLMM 01 Experiment 1 in ESM
```{r echo=FALSE}
round(anova(null, full, test="Chisq"),3)
```

The full model is significantly different than the null model.

##Model output GLMM 01 Experiment 1 in ESM
####Coefficients

```{r echo=FALSE}
round(summary(full)$coefficients, 3)
```

####Individual predictor: Likelihood tests 

```{r echo=FALSE}
xdrop1=drop1(full, test="Chisq",control=contr)
round(xdrop1,3)
```

####Confidence intervals

```{r}
# model1CI=boot.glmm.pred(model.res=full, excl.warnings=T, nboots=1000, para=T)
# round(model1CI$ci.estimates, 3)
# save.image("model1_CIs.RData")
load("model1_CIs.RData")
round(model1CI$ci.estimates, 3)
```


**The interactions terms are not significant (although there is a trend for scores to vary by age) so I will remove them from the model.**

##Reduced model without the interactions - GLMM 02 Experiment 1 in ESM
##Full model 2 

```{r echo=TRUE}
full2=glmer(correct ~ trialtype+order3+z.trialno+z.age+sex+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

##Null model 2

```{r echo=TRUE}
null2<-glmer(correct~ z.trialno+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

####Full and null 2 comparision - GLMM 02 Experiment 1 in ESM

```{r echo=FALSE}
round(anova(null2, full2, test="Chisq"),3)
```

Full model is significantly different than the null model.

##Model output GLMM 02 Experiment 1 in ESM
####Coefficients

```{r echo=FALSE}
round(summary(full2)$coefficients, 3)
```

####Individual predictors : Likelihod ratio tests
  
  
```{r echo=FALSE}
xdrop1=drop1(full2, test="Chisq",control=contr)
round(xdrop1,3)
```

**Age emerges as a significant predictor of scores.**
 
####Confidence intervals for the reduced model without the interaction - GLMM 02 Experiment 1 in ESM

```{r}
# model2CI=boot.glmm.pred(model.res=full2, excl.warnings=T, nboots=1000, para=T)
# round(model2CI$ci.estimates, 3)
# save.image("model2_CIs.RData")
load("model2_CIs.RData")
round(model2CI$ci.estimates, 3)
```

##Scatterplot to see the effect of age on overall performance 

*Running a separate model for the CI calculation.

```{r echo=TRUE}
#full.CI=glmer(correct ~ z.trialno+order3.transfer.c+trialtype.sticker.c+z.age+sex.m.c+(1+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

*Getting confidence intervals for fitted values - bootstrapping

```{r include=FALSE}
#boot.res=boot.glmm.pred(model.res=full.CI, excl.warnings=T, nboots=1000, resol=68, level=0.95, use="z.age", n.cores="all-1", para=T)#with para=T your computer will get very slow so I recommend using this option particularly when you don't need to work with the PC in parallel (the CI calculation will be a lot faster though)

#boot.res$ci.estimates
#boot.res$ci.predicted
#save.image("ripped_foil_child_CIs.RData")
load("ripped_foil_child_CIs.RData")
```

####Overall performance across ages - Figure S2

```{r echo=FALSE}
ripchild_individual <- ripchild %>%
  group_by(id, age, agemonths, openended) %>% 
  summarize(correct = mean(correct))%>%
  ungroup()%>%
  group_by(agemonths)%>%
    add_count(correct)

plot.xvals=seq(from=min(ripchild_individual$agemonths),to=max(ripchild_individual$agemonths),length.out=68)

##using ggplot but using the values from ripchild_individual instead)

scatter<-ggplot(ripchild_individual, aes(x=plot.xvals, y=boot.res$ci.predicted$fitted)) + geom_point(aes(x=agemonths, y=correct), size=ripchild_individual$n, alpha=0.8)+
geom_ribbon(data=boot.res$ci.predicted, aes(x=plot.xvals, ymin=boot.res$ci.predicted$lower.cl,ymax=boot.res$ci.predicted$upper.cl), fill="grey", alpha=0.5)+
geom_line(data=boot.res$ci.predicted, aes(x = plot.xvals, y=boot.res$ci.predicted$fitted), lty=2)+ scale_x_continuous(breaks = seq(from = 36, to = 72, by = 6)) +  geom_segment(aes(x = 42, y = 0, xend = 42, yend = 0.5), arrow=arrow(length = unit(0.2, "cm")), colour="red")+
theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ xlab("Age in months")+ ylab("Mean correct responses")+ 
  ylim(0,1)+
  geom_hline(yintercept=0.50, color='red', linetype="dashed")

scatter
```

**Young children (up until 42 months of age) perform at chance level in this task.**


####Performance in Test and Transfer across ages (w/ responses to the openended question at the end) - Figure 2

####Running a new model for calculating CIs for the phase*age interaction.

```{r include=TRUE}
#full.CI.int<-glmer(correct ~ trialtype.sticker.c+z.age*order3+sex.m.c+z.trialno+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)

#boot.res.int=boot.glmm.pred(model.res=full.CI.int, excl.warnings=T, nboots=1000, resol=136, level=0.95, use="z.age", n.cores="all-1", para=T) #resol=136 this time because data is aggregated for test and transfer separately.
#boot.res.int$ci.estimates
#boot.res.int$ci.predicted

#save.image("ripped_foil_child_CIs_interaction.RData")
load("ripped_foil_child_CIs_interaction.RData")
```


```{r echo=TRUE}
ripchild_separate2 <- ripchild %>%
  group_by(id,  agemonths, order3, openended) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(agemonths, order3)%>%
    add_count(correct)

plot.int=seq(from=min(ripchild_separate2$agemonths),
             to=max(ripchild_separate2$agemonths),length.out=68)

ci.data<-boot.res.int$ci.predicted #I reduced the number of values for the confidence intervals. did not match the actual data.
toDelete<- seq(1, nrow(ci.data),2)
ci.data<-ci.data[toDelete,]


scatter.int3 <- ggplot(data=ripchild_separate2, aes(x= ci.data$z.age, y = ci.data$fitted)) +
  geom_point(data=ripchild_separate2, aes(x = agemonths, y = correct, shape=openended), size = ripchild_separate2$n, alpha = 0.8) +
  facet_wrap(~order3) + geom_segment(data=data.frame(order3="transfer"), mapping=aes(x=44, xend=44, y=0, yend=0.5, order3="transfer"), arrow=arrow(length=unit(0.2,"cm")))+ geom_segment(data=data.frame(order3="test"), mapping=aes(x=40, xend=40, y=0, yend=0.5, order3="test"), arrow=arrow(length=unit(0.2,"cm"))) +
  geom_ribbon(data = ci.data%>%filter(order3=="test"), aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl, group = order3), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data%>%filter(order3=="test"), aes(x = plot.int, y = fitted), lty = 2) +
    geom_ribbon(data = ci.data%>%filter(order3=="transfer"), aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl, group = order3), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data%>%filter(order3=="transfer"), aes(x = plot.int, y = fitted), lty = 2) + scale_shape_manual(name="Explanation", values = c(19,21))+ scale_x_continuous(breaks = seq(from = 36, to = 72, by = 8))+
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Age in months") +
  ylab("Mean correct responses") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.int3
```

```{r echo=TRUE}
####POST-REVIEW Changes - R1 finds the arrows unconventional. Removing them from the figure.

ripchild_separate2 <- ripchild %>%
  group_by(id,  agemonths, order3, openended) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(agemonths, order3)%>%
    add_count(correct)

plot.int=seq(from=min(ripchild_separate2$agemonths),
             to=max(ripchild_separate2$agemonths),length.out=68)

ci.data<-boot.res.int$ci.predicted #I reduced the number of values for the confidence intervals. did not match the actual data.
toDelete<- seq(1, nrow(ci.data),2)
ci.data<-ci.data[toDelete,]


scatter.int3 <- ggplot(data=ripchild_separate2, aes(x= ci.data$z.age, y = ci.data$fitted)) +
  geom_point(data=ripchild_separate2, aes(x = agemonths, y = correct, shape=openended), size = ripchild_separate2$n, alpha = 0.8) +
  facet_wrap(~order3) +
  geom_ribbon(data = ci.data%>%filter(order3=="test"), aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl, group = order3), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data%>%filter(order3=="test"), aes(x = plot.int, y = fitted), lty = 2) +
    geom_ribbon(data = ci.data%>%filter(order3=="transfer"), aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl, group = order3), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data%>%filter(order3=="transfer"), aes(x = plot.int, y = fitted), lty = 2) + scale_shape_manual(name="Explanation", values = c(19,21))+ scale_x_continuous(breaks = seq(from = 36, to = 72, by = 8))+
  theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Age in months") +
  ylab("Mean correct responses") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.int3
ggsave("performance in test and transfer across age w explanation.png",width = 10, height = 5, scale=0.7)
```



##First trial performance in Transfer phase
We expect children to make more errors when they switch from test to transfer if they are solving the task using an arbitrary rule.
*Getting the data for the first trial of Transfer

```{r echo=FALSE}
ripchild_1st_trial <- ripchild %>%
  filter(order3=="transfer" & trialno=="1") %>%
  group_by(id, age, agemonths, sex, trialtype, z.age, trialtype.sticker.c) %>% 
  summarize(correct)
ripchild_1st_trial$z.age=as.vector(scale(ripchild_1st_trial$agemonths))
```

####Performance in trial 1 of Transfer phase - binomial test + GLM

```{r echo=TRUE}
##binomial test for the first trial performance in Transfer
binom.test(x=sum(ripchild_1st_trial$correct), n=length(ripchild_1st_trial$correct), p=0.5, alternative = "two.sided")

firsttrial <- glm(correct~z.age,data=ripchild_1st_trial,family=binomial(link = "logit"))
nullfirsttrial<-glm(correct~1, data=ripchild_1st_trial, family=binomial(link="logit"))
anova(nullfirsttrial, firsttrial, test="Chisq")
summary(firsttrial)$coefficients #Table S4
confint(object=firsttrial)
cbind(coefficients(firsttrial), confint(object=firsttrial))
xdrop1=drop1(firsttrial, test="Chisq",control=contr)
round(xdrop1,3)
```

**Age emerges as a significant predictor of first trial performance of Transfer.** 

```{r echo=FALSE}
#Plotting first trial performance in Transfer
ggplot(ripchild_1st_trial, aes(x=agemonths, y=correct)) + geom_point() + 
  stat_smooth(method="glm", fullrange = TRUE, lty=2, size=0.5, colour="Black", alpha=0.2,  method.args=list(family="binomial")) + scale_x_continuous(breaks = seq(from = 36, to = 72, by = 2)) +  scale_y_continuous(breaks = seq(0, 1, by = 1)) +geom_segment(aes(x = 50, y = 0, xend = 50, yend = 0.5), arrow=arrow(length = unit(0.2, "cm")), colour="red")+ylab("First trial performance of Transfer Phase")+xlab("Age in months")+theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + geom_hline(yintercept=0.50, color='red', linetype="dashed")
ggsave("first trial in transfer.png")
```



```{r echo=FALSE}
####POST-REVIEW Changes - R1 finds the arrows unconventional. Removing them from the figure.
#Plotting first trial performance in Transfer
ggplot(ripchild_1st_trial, aes(x=agemonths, y=correct)) + geom_point() + 
  stat_smooth(method="glm", fullrange = TRUE, lty=2, size=0.5, colour="Black", alpha=0.2,  method.args=list(family="binomial")) + scale_x_continuous(breaks = seq(from = 36, to = 72, by = 2)) +  scale_y_continuous(breaks = seq(0, 1, by = 1)) +ylab("First trial performance of Transfer Phase")+xlab("Age in months")+theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + geom_hline(yintercept=0.50, color='red', linetype="dashed")
ggsave("first trial in transfer.png")
```

**Young children (below 50 months) do not perform differently than chance level** - Figure S3

##Analysis of the open ended questions - Table S5

```{r}
openended<-read.csv("open_ended_ripped.csv", header=T)

openended$z.age=as.vector(scale(openended$agemonths)) #centering age
str(openended)
table(openended$explanation)


oequestion <- glm(explanation~z.age, data=openended, family=binomial(link = "logit"))
nullopen <- glm(explanation ~ 1, data=openended, family=binomial(link="logit"))
anova(nullopen, oequestion, test="Chisq")
summary(oequestion)$coefficients
confint(object=oequestion)
cbind(coefficients(oequestion), confint(object=oequestion))
```






